Exact Phylodynamic Likelihood\\ via Structured Markov Genealogy Processes

### 2024-06-07 ms.Rnw
improvements to examples

<<sirs3>>=
data.frame(
  Beta=4,gamma=2,psi=1,omega=1,
  S0=97,I0=3,R0=0,t0=0,time=40
) -> sirs.params

bake(
  file="sirs3a.rds",
  seed=328168304L,
  dependson=sirs.params,
  {
    library(phylopomp)
    sirs.params |>
      with(
        runSIRS(
          Beta=Beta,gamma=gamma,psi=psi,omega=omega,
          S0=S0,I0=I0,R0=R0,t0=t0,time=time
        )
      )
  }
) -> sirs_tree

bake(
  file="sirs3b.rds",
  seed=621400057L,
  dependson=list(sirs_tree,sirs.params),
  {
    sirs.params |>
      with(
        expand_grid(
          Beta=Beta,
          ##      gamma=gamma,
          gamma=seq(1.7,2.3,by=0.02),
          psi=psi,
          omega=omega,
          ##      omega=seq(0.5,2,by=0.05),
          S0=S0,I0=I0,R0=R0,
          t0=t0,
          rep=1:16,
          Np=5000,
          )
      ) -> params

    foreach (
      p=iter(params,"row"),
      .combine=bind_rows,
      .options.future=list(seed=TRUE)
    ) %dofuture% {
      library(pomp)
      library(phylopomp)
      p |>
        with({
          sirs_tree |>
            sirs_pomp(
              Beta=Beta,gamma=gamma,psi=psi,omega=omega,
              S0=S0,I0=I0,R0=R0,t0=0
            ) |>
            pfilter(Np=Np) |>
            logLik()
        }) -> ll
      bind_cols(p,logLik=ll)
    } -> params
  }
) -> params

@@

\begin{figure}
  \begin{center}
    <<sirs3_plot,dependson="sirs1",fig.dim=c(8,2.8),out.width="100%">>=
    plot_grid(
      A=sirs_tree |>
        plot(points=FALSE,palette=c("#000000"))+
        labs(x="time"),
      B=params |>
        ggplot(aes(x=gamma,y=logLik))+
        geom_point(alpha=0.2)+
        geom_line(data=mcap$fit,aes(x=parameter,y=smoothed),color="blue")+
        geom_vline(xintercept=sirs.params$gamma,color="red")+
        geom_vline(xintercept=mcap$ci,linetype=2)+
        geom_hline(
          yintercept=max(mcap$fit$smoothed)-c(0,mcap$delta),
          linetype=2
        )+
        labs(
          color=character(0),
          y="log likelihood",
          x=expression(gamma)
        )+
        ## lims(y=c(max(params$logLik)-16,NA))+
        theme_classic(),
      labels="AUTO",
      nrow=1,
      rel_widths=c(1,1)
    )
    @
  \end{center}
  \caption{
    \label{fig:sirs_example}
    Likelihood computation for the SIRS model by Sequential Monte Carlo.
    (A)~A~simulated genealogy for $\beta=\Sexpr{sirs.params$Beta}$, $\gamma=\Sexpr{sirs.params$gamma}$, $\omega=\Sexpr{sirs.params$omega}$, $\psi=\Sexpr{sirs.params$psi}$, $(S_0,I_0,R_0)=(\Sexpr{with(sirs.params,c(S0,I0,R0))})$.
    (B)~A~slice through the likelihood surface at the true parameters in the $\gamma$-direction.
    Each point is a distinct Monte Carlo estimate.
    The blue curve is a LOESS smooth;
    the dashed lines bound the Monte Carlo-adjusted 95\% confidence interval \citep{Ionides2017}.
  }
\end{figure}

@@

\begin{table}
  \caption{
    \label{tab:seirs_integ}
    Elements of a scheme for numerically computing the likelihood under the SEIRS model.
    The regular portion of the filter equation holds in between genealogical events (\ie for $t\notin{{B}\cup{S_0}\cup{S_1}}$);
    the singular portion describes the effect of these events ($t\in{{B}\cup{S_0}\cup{S_1}}$).
    Each line corresponds to a potential event, but only those for which $Q=1$ appear in the equation.
    An event with mark $u$ and saturation $s$ has the boost given by the binomial ratio shown (third column).
    The $y\mapsto{y'}$ column depicts the proposed painting schematically, and $\pi$ is the probability of that proposal.
    Blue is used for the $\lab{E}$ deme and yellow for the $\lab{I}$ deme.
    For each line, the filter equation contains $m$ terms.
    An asterisk ($\ast$) stands for cases not explicitly mentioned.
  }
  %% \renewcommand{\arraystretch}{1.2}
  \begin{tabular}{c c c c c c c c | c}
    \hline\hline
    & $u$ & $s$ & $Q$ & $\BinRatio{n}{\ell}{r}{s}$ & $y\mapsto{y'}$ & $\pi$ & $m$ & Line\\[2ex]
    \hline
    \multirow[c]{10}{*}[-1ex]{\rotatebox{90}{$t\notin{{B}\cup{S_0}\cup{S_1}}$}}
    & \multirow{4}{*}{$\lab{Trans}$}
    & $(0,0)$ & 1 & $\left(\frac{E-\ell_E}{E}\right)\left(\frac{I-\ell_I}{I}\right)$ \bigstrut & \unbroken \bigstrut & $\frac{I+1-\ell_I}{I+1}$ & 1 & 1 \\
    \cline{3-8}
    && $(1,0)$ & 1 & $\frac{I-\ell_I}{E I}$ \bigstrut & \maizeblue \bigstrut & $\frac{1}{2(I+1)}$ & $\ell_I$ & 2 \\
    \cline{3-8}
    && $(0,1)$ & 1 & $\frac{E-\ell_E}{E I}$ \bigstrut & \maizemaize \bigstrut & $\frac{1}{2(I+1)}$ & $\ell_I$ & 3 \\
    \cline{3-8}
    && $(1,1)$ & 0 & & & & 0 & 4 \\
    \cline{2-8}
    &\multirow{2}{*}{$\lab{Prog}$}
    & $(0,0)$ & 1 & $\frac{I-\ell_I}{I}\,\Indicator{E\ge\ell_E}$ \bigstrut & \unbroken \bigstrut & $\frac{E+1-\ell_E}{E+1}\,\Indicator{E>\ell_E}$ \bigstrut & 1 & 5 \\
    \cline{3-8}
    && $(0,1)$ & 1 & $\frac{1}{I}\,\Indicator{E\ge\ell_E}$ \bigstrut & \bluemaize \bigstrut & $\frac{1}{E+1}\,\Indicator{E>\ell_E}$ \bigstrut & $\ell_E$ & 6 \\
    \cline{2-8}
    & $\lab{Recov}$ & $(0,0)$ & 1 & $\Indicator{I\ge\ell_I}$ \bigstrut & \unbroken \bigstrut & $\Indicator{I>\ell_I}$ \bigstrut & 1 & 7 \\
    \cline{2-8}
    & $\lab{Wane}$ & $(0,0)$ & 1 & 1 \bigstrut & \unbroken \bigstrut & 1 & 1 & 8 \\
    \cline{2-8}
    & $\lab{Sample}$ & $\ast$ & 0 &  &  & & 0 & 9 \\
    \cline{1-8}
    \multirow[c]{4}{*}[-3ex]{\rotatebox{90}{$t\in{B}$}}
    & \multirow{4}{*}{$\lab{Trans}$}
    & \multirow{4}{*}{$(1,1)$}
    & 1 & $\frac{1}{E I}$ \bigstrut & \branchup \bigstrut & $\frac{1}{2}$ \bigstrut & 1 & 10 \\
    \cline{4-8}
    &&& 1 & $\frac{1}{E I}$ \bigstrut & \branchdown \bigstrut & $\frac{1}{2}$ \bigstrut & 1 & 11 \\
    \cline{4-8}
    &&& 0 &  \bigstrut & \branchupnix \bigstrut & & 0 & 12 \\
    \cline{4-8}
    &&& 0 &  \bigstrut & \branchdownnix \bigstrut & & 0 & 13 \\
    \cline{2-8}
    & $\ast$ & $\ast$ & 0 &  \bigstrut &  & & 0 & 14 \\
    \cline{1-8}
    \multirow[c]{2}{*}[-2ex]{\rotatebox{90}{$t\in{S_0}$} \bigstrut}
    & \multirow{2}{*}{$\lab{Sample}$}
    & \multirow{2}{*}{$(0,0)$}
    & 1 & $\frac{I-\ell_I}{I}$ \bigstrut & \samplezero & 1 & 1 & 15 \\
    \cline{4-8}
    &&& 0 &  & \samplezeronix \bigstrut & & 0 & 16 \\
    \cline{2-8}
    & $\ast$ & $\ast$ & 0 &  &  & & 0 & 17 \\
    \cline{1-8}
    \multirow[c]{2}{*}[-2ex]{\rotatebox{90}{$t\in{S_1}$}}
    & \multirow{2}{*}{$\lab{Sample}$}
    & \multirow{2}{*}{$(0,1)$}
    & 1 & $\frac{1}{I}$ \bigstrut & \sampleone \bigstrut & 1 & 1 & 18 \\
    \cline{4-8}
    &&& 0 &  & \sampleonenix \bigstrut & & 0 & 19 \\
    \cline{2-8}
    & $\ast$ & $\ast$ & 0 &  &  & & 0 & 20 \\
    \hline\hline
  \end{tabular}
\end{table}

@@

<<seirs3>>=
seirs.params <- data.frame(
  Beta=4,sigma=1,gamma=1,psi=1,omega=1,
  S0=200,E0=3,I0=5,R0=100,
  time=3
)

bake(
  file="seirs3a.rds",
  seed=831282841L,
  dependson=seirs.params,
  seirs.params |>
    with(
      runSEIR(
        Beta=Beta,sigma=sigma,gamma=gamma,psi=psi,omega=omega,
        S0=S0,E0=E0,I0=I0,R0=R0,
        time=time
      )
    )
) -> seirs_tree

bake(
  file="seirs3b.rds",
  seed=831282841L,
  dependson=list(seirs.params,seirs_tree),
  {
    library(phylopomp)
    library(circumstance)
    library(doFuture)
    plan(multicore)

    seirs.params |>
      with(
        seirs_tree |>
          seirs_pomp(
            Beta=Beta,sigma=sigma,gamma=gamma,psi=psi,omega=omega,
            S0=S0,E0=E0,I0=I0,R0=R0
          )
      ) -> po

    seq(0.2,2.5,by=0.1) |>
      lapply(
        \(s) {
          x <- po
          coef(x,"sigma") <- s
          x
        }
      ) |>
      concat() |>
      pfilter(Np=20000,Nrep=10)
  }
) -> pfs

@@

\begin{figure}
  \begin{center}
    <<seirs3_plot,fig.dim=c(8,2.8),out.width="100%",dependon="seirs3">>=
    plot_grid(
      A=seirs_tree |>
        plot(points=TRUE,palette="#000000")+
        labs(x="time"),
      B=params |>
        ggplot()+
        geom_point(aes(x=sigma,y=logLik))+
        geom_line(data=mcap$fit,aes(x=parameter,y=smoothed),color="blue")+
        geom_vline(xintercept=seirs.params$sigma,color="red")+
        geom_vline(xintercept=mcap$ci,linetype=2)+
        geom_hline(
          yintercept=max(mcap$fit$smoothed)-c(0,mcap$delta),
          linetype=2
        )+
        ##        lims(y=c(max(params$logLik)-12,NA))+
        labs(
          color=character(0),
          y="log likelihood",
          x=expression(sigma)
        )+
        theme_classic(),
      labels="AUTO",
      nrow=1,
      rel_widths=c(3,4)
    )
    @
  \end{center}
  \caption{
    \label{fig:seirs_example}
    Likelihood computation for the SEIRS model by Sequential Monte Carlo, using the scheme of \cref{box:SEIRS}.
    (A)~Simulated genealogy for $\beta=\Sexpr{seirs.params$Beta}$,
    $\sigma=\Sexpr{seirs.params$sigma}$,
    $\gamma=\Sexpr{seirs.params$gamma}$,
    $\psi=\Sexpr{seirs.params$psi}$,
    $\omega=\Sexpr{seirs.params$omega}$,
    $(S_0,E_0,I_0,R_0)=(\Sexpr{with(seirs.params,c(S0,E0,I0,R0))})$.
    (B)~Likelihood slice in the $\sigma$-direction.
    Each point represents the estimate of an independent Sequential Monte Carlo computation.
    The blue curve shows a LOESS smooth;
    the dashed vertical lines enclose the Monte Carlo-adjusted 95\% confidence interval \citep{Ionides2017}.
  }
\end{figure}

>>>>

<<sirs3>>=
data.frame(
  Beta=4,gamma=2,psi=1,omega=1,
  S0=97,I0=3,R0=0,t0=0,time=40
) -> sirs_params

bake(
  file="sirs3a.rds",
  seed=328168304L,
  dependson=sirs_params,
  {
    library(phylopomp)
    sirs_params |>
      with(
        runSIRS(
          Beta=Beta,gamma=gamma,psi=psi,omega=omega,
          S0=S0,I0=I0,R0=R0,t0=t0,time=time
        )
      )
  }
) -> sirs_tree

sirs_params |>
  with(
    expand_grid(
      Beta=Beta,
      gamma=seq(1.7,2.3,by=0.02),
      psi=psi,
      omega=omega,
      S0=S0,I0=I0,R0=R0,
      t0=t0,
      rep=seq_len(8),
      Np=10000,
      )
  ) -> params

bake(
  file="sirs3b.rds",
  seed=621400057L,
  dependson=list(sirs_tree,sirs_params,params),
  {
    library(iterators)
    library(doFuture)
    plan(multicore)

    foreach (
      p=iter(params,"row")
    ) %dofuture% {
      library(pomp)
      library(phylopomp)
      p |>
        with({
          sirs_tree |>
            sirs_pomp(
              Beta=Beta,gamma=gamma,psi=psi,omega=omega,
              S0=S0,I0=I0,R0=R0,t0=0
            ) |>
            pfilter(Np=Np)
        })
    } %seed% TRUE |>
      concat()
  }) -> pfs

@@

\begin{figure}
  \begin{center}
    <<sirs3_plot,dependson="sirs1",fig.dim=c(8,2.8),out.width="100%">>=
    plot_grid(
      A=sirs_tree |>
        plot(points=FALSE,palette=c("#000000"))+
        labs(x="time"),
      B=params |>
        ggplot(aes(x=gamma,y=logLik))+
        geom_point(alpha=0.4)+
        geom_line(data=mcap$fit,aes(x=parameter,y=smoothed),color="blue")+
        geom_vline(xintercept=sirs_params$gamma,color="red")+
        geom_vline(xintercept=mcap$ci,linetype=2)+
        geom_hline(
          yintercept=max(mcap$fit$smoothed)-c(0,mcap$delta),
          linetype=2
        )+
        labs(
          color=character(0),
          y="log likelihood",
          x=expression(gamma)
        )+
        lims(y=c(max(params$logLik)-12,NA))+
        theme_classic(),
      labels="AUTO",
      nrow=1,
      rel_widths=c(1,1)
    )
    @
  \end{center}
  \caption{
    \label{fig:sirs_example}
    Likelihood computation for the SIRS model by Sequential Monte Carlo.
    (A)~A~simulated genealogy for $\beta=\Sexpr{sirs_params$Beta}$, $\gamma=\Sexpr{sirs_params$gamma}$, $\omega=\Sexpr{sirs_params$omega}$, $\psi=\Sexpr{sirs_params$psi}$, $(S_0,I_0,R_0)=(\Sexpr{with(sirs_params,c(S0,I0,R0))})$.
    (B)~A~slice through the likelihood surface at the true parameters in the $\gamma$-direction.
    Each point is a distinct Monte Carlo estimate.
    The blue curve is a LOESS smooth;
    the dashed lines bound the Monte Carlo-adjusted 95\% confidence interval \citep{Ionides2017}.
  }
\end{figure}

@@

\begin{table}
  \caption{
    \label{tab:seirs_integ}
    Elements of a scheme for numerically computing the likelihood under the SEIRS model.
    The regular portion of the filter equation holds in between genealogical events (\ie for $t\notin{{B}\cup{S_0}\cup{S_1}}$);
    the singular portion describes the effect of these events ($t\in{{B}\cup{S_0}\cup{S_1}}$).
    Each line corresponds to a potential event, but only those for which $Q=1$ appear in the equation.
    An event with mark $u$ and saturation $s$ has the boost given by the binomial ratio shown (third column).
    The $y\mapsto{y'}$ column depicts the proposed painting schematically, and $\pi$ is the probability of that proposal.
    Blue is used for the $\lab{E}$ deme and yellow for the $\lab{I}$ deme.
    For each line, the filter equation contains $m$ terms.
    An asterisk ($\ast$) stands for cases not explicitly mentioned.
  }
  %% \renewcommand{\arraystretch}{1.2}
  \begin{tabular}{c c c c c c c c | c}
    \hline\hline
    & $u$ & $s$ & $Q$ & $\BinRatio{n}{\ell}{r}{s}$ & $y\mapsto{y'}$ & $\pi$ & $m$ & Line\\[2ex]
    \hline
    \multirow[c]{10}{*}[-1ex]{\rotatebox{90}{$t\notin{{B}\cup{S_0}\cup{S_1}}$}}
    & \multirow{4}{*}{$\lab{Trans}$}
    & $(0,0)$ & 1 & $\left(\frac{E-\ell_E}{E}\right)\left(\frac{I-\ell_I}{I}\right)$ \bigstrut & \unbroken \bigstrut & $\frac{I-\ell_I}{I}$ & 1 & 1 \\
    \cline{3-8}
    && $(1,0)$ & 1 & $\frac{I-\ell_I}{E I}$ \bigstrut & \maizeblue \bigstrut & $\frac{1}{2I}$ & $\ell_I$ & 2 \\
    \cline{3-8}
    && $(0,1)$ & 1 & $\frac{E-\ell_E}{E I}$ \bigstrut & \maizemaize \bigstrut & $\frac{1}{2I}$ & $\ell_I$ & 3 \\
    \cline{3-8}
    && $(1,1)$ & 0 & & & & 0 & 4 \\
    \cline{2-8}
    &\multirow{2}{*}{$\lab{Prog}$}
    & $(0,0)$ & 1 & $\frac{I-\ell_I}{I}\,\Indicator{E\ge\ell_E}$ \bigstrut & \unbroken \bigstrut & $\frac{E-\ell_E}{E}\,\Indicator{E>\ell_E}$ \bigstrut & 1 & 5 \\
    \cline{3-8}
    && $(0,1)$ & 1 & $\frac{1}{I}\,\Indicator{E\ge\ell_E}$ \bigstrut & \bluemaize \bigstrut & $\frac{1}{E}\,\Indicator{E>\ell_E}$ \bigstrut & $\ell_E$ & 6 \\
    \cline{2-8}
    & $\lab{Recov}$ & $(0,0)$ & 1 & $\Indicator{I\ge\ell_I}$ \bigstrut & \unbroken \bigstrut & $\Indicator{I>\ell_I}$ \bigstrut & 1 & 7 \\
    \cline{2-8}
    & $\lab{Wane}$ & $(0,0)$ & 1 & 1 \bigstrut & \unbroken \bigstrut & 1 & 1 & 8 \\
    \cline{2-8}
    & $\lab{Sample}$ & $\ast$ & 0 &  &  & & 0 & 9 \\
    \cline{1-8}
    \multirow[c]{4}{*}[-3ex]{\rotatebox{90}{$t\in{B}$}}
    & \multirow{4}{*}{$\lab{Trans}$}
    & \multirow{4}{*}{$(1,1)$}
    & 1 & $\frac{1}{E I}$ \bigstrut & \branchup \bigstrut & $\frac{1}{2}$ \bigstrut & 1 & 10 \\
    \cline{4-8}
    &&& 1 & $\frac{1}{E I}$ \bigstrut & \branchdown \bigstrut & $\frac{1}{2}$ \bigstrut & 1 & 11 \\
    \cline{4-8}
    &&& 0 &  \bigstrut & \branchupnix \bigstrut & & 0 & 12 \\
    \cline{4-8}
    &&& 0 &  \bigstrut & \branchdownnix \bigstrut & & 0 & 13 \\
    \cline{2-8}
    & $\ast$ & $\ast$ & 0 &  \bigstrut &  & & 0 & 14 \\
    \cline{1-8}
    \multirow[c]{2}{*}[-2ex]{\rotatebox{90}{$t\in{S_0}$} \bigstrut}
    & \multirow{2}{*}{$\lab{Sample}$}
    & \multirow{2}{*}{$(0,0)$}
    & 1 & $\frac{I-\ell_I}{I}$ \bigstrut & \samplezero & 1 & 1 & 15 \\
    \cline{4-8}
    &&& 0 &  & \samplezeronix \bigstrut & & 0 & 16 \\
    \cline{2-8}
    & $\ast$ & $\ast$ & 0 &  &  & & 0 & 17 \\
    \cline{1-8}
    \multirow[c]{2}{*}[-2ex]{\rotatebox{90}{$t\in{S_1}$}}
    & \multirow{2}{*}{$\lab{Sample}$}
    & \multirow{2}{*}{$(0,1)$}
    & 1 & $\frac{1}{I}$ \bigstrut & \sampleone \bigstrut & 1 & 1 & 18 \\
    \cline{4-8}
    &&& 0 &  & \sampleonenix \bigstrut & & 0 & 19 \\
    \cline{2-8}
    & $\ast$ & $\ast$ & 0 &  &  & & 0 & 20 \\
    \hline\hline
  \end{tabular}
\end{table}

@@

<<seirs3>>=
seirs_params <- data.frame(
  Beta=3,sigma=1,gamma=0.5,psi=0.02,omega=0.08,
  S0=70,E0=1,I0=0,R0=50,
  time=400
)

bake(
  file="seirs3a.rds",
  dependson=seirs_params,
  seed=509673338,
  seirs_params |>
    with(
      runSEIR(
        Beta=Beta,sigma=sigma,gamma=gamma,psi=psi,omega=omega,
        S0=S0,E0=E0,I0=I0,R0=R0,
        time=time
      )
    )
)-> seirs_tree

seirs_params |>
  select(-sigma,-time) |>
  expand_grid(
    sigma=seq(0.4,2.4,length.out=25),
    rep=seq_len(8)
  ) |>
  mutate(N=S0+E0+I0+R0) |>
  collect() -> params

bake(
  file="seirs3b.rds",
  dependson=list(params,seirs_params,seirs_tree),
  seed=751601556,
  {
    seirs_params |>
      with(
        seirs_tree |>
          seirs_pomp(
            Beta=Beta,sigma=sigma,gamma=gamma,psi=psi,omega=omega,
            S0=S0,E0=E0,I0=I0,R0=R0
          )
      ) -> po

    library(iterators)
    library(doFuture)
    plan(multicore)
    foreach (
      p=iter(params,"row")
    ) %dofuture% {
      library(phylopomp)
      po |>
        pfilter(params=p,Np=1e4)
    } %seed% TRUE |>
      concat()
  }
) -> pfs

@@

\begin{figure}
  \begin{center}
    <<seirs3_plot,fig.dim=c(8,2.8),out.width="100%",dependon="seirs3">>=
    plot_grid(
      A=seirs_tree |>
        plot(points=FALSE,palette="#000000")+
        labs(x="time"),
      B=params |>
        ggplot()+
        geom_point(aes(x=sigma,y=logLik))+
        geom_line(data=mcap$fit,aes(x=parameter,y=smoothed),color="blue")+
        geom_vline(xintercept=seirs_params$sigma,color="red")+
        geom_vline(xintercept=mcap$ci,linetype=2)+
        geom_hline(
          yintercept=with(mcap,max(fit$smoothed)-c(0,delta)),
          linetype=2
        )+
        lims(y=c(max(params$logLik)-12,NA))+
        labs(
          color=character(0),
          y="log likelihood",
          x=expression(sigma)
        )+
        theme_classic(),
      labels="AUTO",
      nrow=1,
      rel_widths=c(3,4)
    )
    @
  \end{center}
  \caption{
    \label{fig:seirs_example}
    Likelihood computation for the SEIRS model by Sequential Monte Carlo, using the scheme of \cref{box:SEIRS}.
    (A)~Simulated genealogy for $\beta=\Sexpr{seirs_params$Beta}$,
    $\sigma=\Sexpr{seirs_params$sigma}$,
    $\gamma=\Sexpr{seirs_params$gamma}$,
    $\psi=\Sexpr{seirs_params$psi}$,
    $\omega=\Sexpr{seirs_params$omega}$,
    $(S_0,E_0,I_0,R_0)=(\Sexpr{with(seirs_params,c(S0,E0,I0,R0))})$.
    (B)~Likelihood slice in the $\sigma$-direction.
    Each point represents the estimate of an independent Sequential Monte Carlo computation.
    The blue curve shows a LOESS smooth;
    the dashed vertical lines enclose the Monte Carlo-adjusted 95\% confidence interval \citep{Ionides2017}.
  }
\end{figure}
